---
title: "Statistical Analysis of Deprivation and ED Referral Source RIE 2022 + 2023"
output:
  html_document:
    df_print: paged
---

Load relevant packages
```{r}
library(readr)
library (finalfit)
library(tidyverse)
```

Read in data 
```{r}
AEdatadeprivation_ReferralSource <- read_csv("/Users/Alex.M.Wrighton/Documents/ED_Deprivation_Project_Alix_Wrighton/AEdatadeprivation_ReferralSource.zip")
View(AEdatadeprivation_ReferralSource)
```

Sort data to include on RIE presentations between 2022 and 2022 for patients age >= 16

```{r}
Data_2022_2023_RIE <- AEdatadeprivation_ReferralSource %>% select (ARRIVAL_DATE, SEX, DEPRV_SCT_QUINTILE, DEPRV_SCT_DECILE, AGE_ON_ARRIVAL, LOCATION_CODE, SOURCE_OF_REFERRAL) %>% filter(LOCATION_CODE == "RIE") %>% filter(ARRIVAL_DATE < "2024-01-01") %>% filter (ARRIVAL_DATE > "2021-12-31") %>%  filter(AGE_ON_ARRIVAL > 15)
```

Remove NAs and unknown sexes

```{r}
Complete_data <- Data_2022_2023_RIE %>% filter(!is.na(ARRIVAL_DATE)) %>% filter(!is.na(SEX)) %>% filter(!is.na(DEPRV_SCT_DECILE)) %>% filter(!is.na(DEPRV_SCT_QUINTILE)) %>% filter(!is.na(AGE_ON_ARRIVAL)) %>% filter(!is.na (LOCATION_CODE)) %>% filter(!is.na (SOURCE_OF_REFERRAL)) %>% filter(SEX %in% c("MALE", "FEMALE"))
```

Check data to ensure no outliers

```{r}
Complete_data %>% count(SEX)
Complete_data %>% count(LOCATION_CODE)
Complete_data %>% count(SOURCE_OF_REFERRAL)
Complete_data %>% ggplot (aes (x= AGE_ON_ARRIVAL)) + geom_histogram()
```

Only include ages <120 years old (2 x patients with age of 120)
```{r}
Complete_data <- Complete_data %>% filter (AGE_ON_ARRIVAL < 120)
```

Changes variables into factors with levels

```{r}
factordata <- Complete_data %>% mutate (sex.factor = factor(SEX) %>% fct_recode ("Male" = "MALE", "Female" = "FEMALE") %>% ff_label ("Sex"))

factordata <- factordata %>% mutate (deprv.factor = factor(DEPRV_SCT_DECILE) %>% fct_recode ("Decile 1" = "1", "Decile 2" = "2", "Decile 3" = "3","Decile 4" = "4", "Decile 5" = "5","Decile 6" = "6", "Decile 7" = "7", "Decile 8" = "8", "Decile 9" = "9", "Decile 10" = "10") %>% ff_label ("Deprivation Decile"))

factordata <- factordata %>% mutate (referral.factor = factor(SOURCE_OF_REFERRAL) %>% fct_recode ("Self-referral" = "Self Referral to A&E", "NHS 24 Referral" = "NHS 24", "Flow Centre Referral" = "Flow Centre", "999 Emergency" = "999 Emergency", "GP referral" = "General Practitioner", "LUCS Referral" = "Lothian Unscheduled Care Services", "Police Referral" = "Police", "Reattend" = "Reattend", "Misc Referral" = "Social Work Dept" , "Misc Referral" = "Leith Community Treatment Centre", "Misc Referral" = "SAS Paramedic Practitioner Referral", "Misc Referral" = "Ophthalmic Referral", "Misc Referral" = "Other", "Misc Referral" = "RVH", "Misc Referral" = "RHCYP/RHSC", "Misc Referral" = "Liberton", "Misc Referral" = "Roodlands", "Misc Referral" = "Astley Ainslie Hospital", "Misc Referral" = "Hospital (Outwith LUHD)", "Misc Referral" = "WGH - MIU", "Misc Referral" = "St Johns", "Misc Referral" = "RIE" ))

factordata <- factordata %>% mutate(age.factor = AGE_ON_ARRIVAL %>% cut(breaks = c(16,30,50,70,120), include.lowest = TRUE) %>% fct_recode("≤30" ="[16,30]","31 to 50" = "(30,50]", "51 to 70" = "(50,70]", ">70"="(70,120]") %>%ff_label("Age (years)"))

```


# GP Referral

Create summarised column for referral by GP or not
```{r}
factordata <- factordata %>% mutate(referral_gp = fct_collapse(referral.factor,"Not referred by GP" = c("999 Emergency", "Flow Centre Referral", "LUCS Referral", "NHS 24 Referral", "Police Referral", "Reattend", "Self-referral", "Misc Referral"))) %>% ff_label ("GP Referral")
```

Create final fit table for GP referral with p-values displayed. Note: Decile 1 removed as results are being compared relative to it

```{r}
factordata %>% summary_factorlist(dependent   = "referral_gp", explanatory = "deprv.factor", p = TRUE)
```

Create model for GP referral vs deprivation + age + sex
```{r}
factordata %>%  summary_factorlist(dependent   = "referral_gp", explanatory = c("deprv.factor", "age.factor", "sex.factor"), p = TRUE)
```


Create model for GP referral vs deprivation
```{r}
fit_gp <- glm(referral_gp ~ deprv.factor, data = factordata, family = binomial)
summary(fit_gp)
library(broom)
fit_gp %>% tidy (conf.int = TRUE, exp = TRUE)
```

Note: Desirable model with low AIC and high c-statistic
Model for GP referral vs deprivation
```{r}
dependent_gp <- "referral_gp"
explanatory_deprv <- "deprv.factor"
```
```{r}
factordata %>% finalfit(dependent_gp, explanatory_deprv, metrics = TRUE)
```


Model for GP referral vs deprivation + age

```{r}
explanatory_deprv_age <- c ("deprv.factor", "age.factor")
factordata %>% finalfit(dependent_gp, explanatory_deprv_age, metrics = TRUE)
```

Model for GP referral vs deprivation + sex

```{r}
explanatory_deprv_sex <- c ("deprv.factor", "sex.factor")
factordata %>% finalfit(dependent_gp, explanatory_deprv_sex, metrics = TRUE)
```

Model for GP referral vs deprivation + age + sex

```{r}
explanatory_deprv_sex_age <- c ("deprv.factor", "age.factor", "sex.factor")
factordata %>% finalfit (dependent_gp, explanatory_deprv_sex_age, metrics = TRUE)
```

Model for GP referral vs age + sex vs deprv+age+sex
```{r}
explanatory_multi <- c ("age.factor", "sex.factor")
factordata %>% finalfit (dependent_gp, explanatory_deprv_sex_age, explanatory_multi, keep_models = TRUE, metrics = TRUE)
```

Model for GP referral vs deprivation +age + sex +age:sex (interaction term)
```{r}
explanatory_allterms <- c ("deprv.factor", "age.factor", "sex.factor", "sex.factor*age.factor")
```
```{r}
factordata %>% finalfit(dependent_gp, explanatory_allterms, metrics = TRUE)
```


Model for GP referral vs age + sex + age:sex
```{r}
explanatory_reduced <- c ("age.factor", "sex.factor", "sex.factor*age.factor")
factordata %>% finalfit(dependent_gp, explanatory_reduced, metrics = TRUE)
```
 
 Likelihood ratio test for GP
```{r}
install.packages("lmtest")
model_fit_gp_deprv <- glm(referral_gp ~ deprv.factor, data = factordata, family = binomial)
model_fit_gp_allterms <- glm(referral_gp ~ deprv.factor + age.factor + sex.factor + age.factor:sex.factor, data = factordata, family = binomial)
lmtest::lrtest(model_fit_gp_deprv, model_fit_gp_allterms)
```
Forest plot for GP
```{r}
tiff("GP_OR_plot.tiff", width = 8, height = 5, units = 'in', res = 300)
GP_ORplot <- factordata %>% or_plot(dependent_gp, explanatory_allterms)
dev.off()
```


# Police Referral

Create summarised column for referral by Police or not
```{r}
factordata <- factordata %>% mutate(referral_police = fct_collapse(referral.factor,"Not referred by Police" = c("999 Emergency", "Flow Centre Referral", "LUCS Referral", "NHS 24 Referral", "GP referral", "Reattend", "Self-referral", "Misc Referral"))) %>% ff_label ("Police Referral")
```

Create final fit table for Police referral with p-values displayed. Note: Decile 1 removed as results are being compared relative to it

```{r}
factordata %>% summary_factorlist(dependent   = "referral_police", explanatory = "deprv.factor", p = TRUE)
```

Create model for Police referral vs deprivation + age + sex
```{r}
factordata %>%  summary_factorlist(dependent   = "referral_police", explanatory = c("deprv.factor", "age.factor", "sex.factor"), p = TRUE)
```


Create model for Police referral vs deprivation
```{r}
fit_police <- glm(referral_police ~ deprv.factor, data = factordata, family = binomial)
summary(fit_police)
library(broom)
fit_police %>% tidy (conf.int = TRUE, exp = TRUE)
```

Note: Desirable model with low AIC and high c-statistic

Model for Police referral vs deprivation
```{r}
dependent_police <- "referral_police"
explanatory_deprv <- "deprv.factor"
```
```{r}
factordata %>% finalfit(dependent_police, explanatory_deprv, metrics = TRUE)
```


Model for Police referral vs deprivation + age

```{r}
explanatory_deprv_age <- c ("deprv.factor", "age.factor")
factordata %>% finalfit(dependent_police, explanatory_deprv_age, metrics = TRUE)
```

Model for Police referral vs deprivation + sex

```{r}
explanatory_deprv_sex <- c ("deprv.factor", "sex.factor")
factordata %>% finalfit(dependent_police, explanatory_deprv_sex, metrics = TRUE)
```

Model for Police referral vs deprivation + age + sex

```{r}
explanatory_deprv_sex_age <- c ("deprv.factor", "age.factor", "sex.factor")
factordata %>% finalfit (dependent_police, explanatory_deprv_sex_age, metrics = TRUE)
```

Model for Police referral vs age + sex vs deprv+age+sex
```{r}
explanatory_multi <- c ("age.factor", "sex.factor")
factordata %>% finalfit (dependent_police, explanatory_deprv_sex_age, explanatory_multi, keep_models = TRUE, metrics = TRUE)
```

Model for Police referral vs deprivation +age + sex +age:sex (interaction term)
```{r}
explanatory_allterms <- c ("deprv.factor", "age.factor", "sex.factor", "sex.factor*age.factor")
```
```{r}
factordata %>% finalfit(dependent_police, explanatory_allterms, metrics = TRUE)
```

Model for Police referral vs age + sex + age:sex
```{r}
explanatory_reduced <- c ("age.factor", "sex.factor", "sex.factor*age.factor")
factordata %>% finalfit(dependent_police, explanatory_reduced, metrics = TRUE)
```
 
 Likelihood ratio test for Police
```{r}
install.packages("lmtest")
model_fit_police_deprv <- glm(referral_police ~ deprv.factor, data = factordata, family = binomial)
model_fit_police_allterms <- glm(referral_police ~ deprv.factor + age.factor + sex.factor + age.factor:sex.factor, data = factordata, family = binomial)
lmtest::lrtest(model_fit_police_deprv, model_fit_police_allterms)
```
Forest plot for Police
```{r}
tiff("Police_OR_plot.tiff", width = 8, height = 5, units = 'in', res = 300)
Police_ORplot <- factordata %>% or_plot(dependent_police, explanatory_allterms)
dev.off()
```

# NHS 24 Referral

Create summarised column for referral by NHS 24 or not
```{r}
factordata <- factordata %>% mutate(referral_nhs24 = fct_collapse(referral.factor,"Not referred by NHS 24" = c("999 Emergency", "Flow Centre Referral", "LUCS Referral", "GP referral", "Police Referral", "Reattend", "Self-referral", "Misc Referral"))) %>% ff_label ("NHS 24 Referral")
```

Create final fit table for NHS 24 referral with p-values displayed. Note: Decile 1 removed as results are being compared relative to it

```{r}
factordata %>% summary_factorlist(dependent   = "referral_nhs24", explanatory = "deprv.factor", p = TRUE)
```

Create model for NHS 24 referral vs deprivation + age + sex
```{r}
factordata %>%  summary_factorlist(dependent   = "referral_nhs24", explanatory = c("deprv.factor", "age.factor", "sex.factor"), p = TRUE)
```


Create model for NHS 24 referral vs deprivation
```{r}
fit_nhs24 <- glm(referral_nhs24 ~ deprv.factor, data = factordata, family = binomial)
summary(fit_nhs24)
library(broom)
fit_nhs24 %>% tidy (conf.int = TRUE, exp = TRUE)

```

Note: Desirable model with low AIC and high c-statistic

Model for NHS 24 referral vs deprivation
```{r}
dependent_nhs24 <- "referral_nhs24"
explanatory_deprv <- "deprv.factor"
```
```{r}
factordata %>% finalfit(dependent_nhs24, explanatory_deprv, metrics = TRUE)
```

Model for NHS 24 referral vs deprivation + age

```{r}
explanatory_deprv_age <- c ("deprv.factor", "age.factor")
factordata %>% finalfit(dependent_nhs24, explanatory_deprv_age, metrics = TRUE)
```

Model for NHS 24 referral vs deprivation + sex

```{r}
explanatory_deprv_sex <- c ("deprv.factor", "sex.factor")
factordata %>% finalfit(dependent_nhs24, explanatory_deprv_sex, metrics = TRUE)
```

Model for NHS 24 referral vs deprivation + age + sex

```{r}
explanatory_deprv_sex_age <- c ("deprv.factor", "age.factor", "sex.factor")
factordata %>% finalfit (dependent_nhs24, explanatory_deprv_sex_age, metrics = TRUE)
```

Model for NHS 24 referral vs age + sex vs deprv+age+sex
```{r}
explanatory_multi <- c ("age.factor", "sex.factor")
factordata %>% finalfit (dependent_nhs24, explanatory_deprv_sex_age, explanatory_multi, keep_models = TRUE, metrics = TRUE)
```

Model for NHS24 referral vs deprivation +age + sex +age:sex (interaction term)
```{r}
explanatory_allterms <- c ("deprv.factor", "age.factor", "sex.factor", "sex.factor*age.factor")
```
```{r}
factordata %>% finalfit(dependent_nhs24, explanatory_allterms, metrics = TRUE)
```

Model for NHS 24 referral vs age + sex + age:sex
```{r}
explanatory_reduced <- c ("age.factor", "sex.factor", "sex.factor*age.factor")
factordata %>% finalfit(dependent_nhs24, explanatory_reduced, metrics = TRUE)
```
 
 Likelihood ratio test for NHS24
```{r}
install.packages("lmtest")
model_fit_nhs24_deprv <- glm(referral_nhs24 ~ deprv.factor, data = factordata, family = binomial)
model_fit_nhs24_allterms <- glm(referral_nhs24 ~ deprv.factor + age.factor + sex.factor + age.factor:sex.factor, data = factordata, family = binomial)
lmtest::lrtest(model_fit_nhs24_deprv, model_fit_nhs24_allterms)
```
Forest plot for NHS24
```{r}
tiff("NHS24_OR_plot.tiff", width = 8, height = 5, units = 'in', res = 300)
NHS24_ORplot <- factordata %>% or_plot(dependent_nhs24, explanatory_allterms)
dev.off()
```

# 999 Emergency Referral

Create summarised column for referral by 999 or not
```{r}
factordata <- factordata %>% mutate(referral_999 = fct_collapse(referral.factor,"Not referred by 999 Emergency" = c("NHS 24 Referral", "Flow Centre Referral", "LUCS Referral", "GP referral", "Police Referral", "Reattend", "Self-referral", "Misc Referral"))) %>% ff_label ("999 Emergency")
```

Create final fit table for 999 referral with p-values displayed. Note: Decile 1 removed as results are being compared relative to it

```{r}
factordata %>% summary_factorlist(dependent   = "referral_999", explanatory = "deprv.factor", p = TRUE)
```

Create model for 999 referral vs deprivation + age + sex
```{r}
factordata %>%  summary_factorlist(dependent   = "referral_999", explanatory = c("deprv.factor", "age.factor", "sex.factor"), p = TRUE)
```


Create model for 999 referral vs deprivation
```{r}
fit_999 <- glm(referral_999 ~ deprv.factor, data = factordata, family = binomial)
summary(fit_999)
library(broom)
fit_999 %>% tidy (conf.int = TRUE, exp = TRUE)
```

Note: Desirable model with low AIC and high c-statistic

Model for 999 referral vs deprivation
```{r}
dependent_999 <- "referral_999"
explanatory_deprv <- "deprv.factor"
```
```{r}
factordata %>% finalfit(dependent_999, explanatory_deprv, metrics = TRUE)
```

Model for 999 referral vs deprivation + age

```{r}
explanatory_deprv_age <- c ("deprv.factor", "age.factor")
factordata %>% finalfit(dependent_999, explanatory_deprv_age, metrics = TRUE)
```

Model for 999 referral vs deprivation + sex

```{r}
explanatory_deprv_sex <- c ("deprv.factor", "sex.factor")
factordata %>% finalfit(dependent_999, explanatory_deprv_sex, metrics = TRUE)
```

Model for 999 referral vs deprivation + age + sex

```{r}
explanatory_deprv_sex_age <- c ("deprv.factor", "age.factor", "sex.factor")
factordata %>% finalfit (dependent_999, explanatory_deprv_sex_age, metrics = TRUE)
```

Model for 999 referral vs age + sex vs deprv+age+sex
```{r}
explanatory_multi <- c ("age.factor", "sex.factor")
factordata %>% finalfit (dependent_999, explanatory_deprv_sex_age, explanatory_multi, keep_models = TRUE, metrics = TRUE)
```

Model for 999 referral vs deprivation +age + sex +age:sex (interaction term)
```{r}
explanatory_allterms <- c ("deprv.factor", "age.factor", "sex.factor", "sex.factor*age.factor")
```
```{r}
factordata %>% finalfit(dependent_999, explanatory_allterms, metrics = TRUE)
```


Model for 999 referral vs age + sex + age:sex
```{r}
explanatory_reduced <- c ("age.factor", "sex.factor", "sex.factor*age.factor")
factordata %>% finalfit(dependent_999, explanatory_reduced, metrics = TRUE)
```
 
 Likelihood ratio test for 999
```{r}
install.packages("lmtest")
model_fit_999_deprv <- glm(referral_999 ~ deprv.factor, data = factordata, family = binomial)
model_fit_999_allterms <- glm(referral_999 ~ deprv.factor + age.factor + sex.factor + age.factor:sex.factor, data = factordata, family = binomial)
lmtest::lrtest(model_fit_999_deprv, model_fit_999_allterms)
```
Forest plot for 999
```{r}

tiff("999_OR_plot.tiff", width = 8, height = 5, units = 'in', res = 300)
Emergency999_ORplot <- factordata %>% or_plot(dependent_999, explanatory_allterms)
dev.off()
```


# Self Referral

Create summarised column for referral by self referral or not
```{r}
factordata <- factordata %>% mutate(referral_selfreferral = fct_collapse(referral.factor,"Not referred by Self Referral" = c("NHS 24 Referral", "Flow Centre Referral", "LUCS Referral", "GP referral", "Police Referral", "Reattend", "999 Emergency", "Misc Referral"))) %>% ff_label ("Self Referral")
```

Create final fit table for self referral with p-values displayed. Note: Decile 1 removed as results are being compared relative to it

```{r}
factordata %>% summary_factorlist(dependent   = "referral_selfreferral", explanatory = "deprv.factor", p = TRUE)
```

Create model for self referral vs deprivation + age + sex
```{r}
factordata %>%  summary_factorlist(dependent   = "referral_selfreferral", explanatory = c("deprv.factor", "age.factor", "sex.factor"), p = TRUE)
```


Create model for self referral vs deprivation
```{r}
fit_selfreferral <- glm(referral_selfreferral ~ deprv.factor, data = factordata, family = binomial)
summary(fit_selfreferral)
library(broom)
fit_selfreferral %>% tidy (conf.int = TRUE, exp = TRUE)
```

Note: Desirable model with low AIC and high c-statistic

Model for self referral vs deprivation
```{r}
dependent_selfreferral<- "referral_selfreferral"
explanatory_deprv <- "deprv.factor"
```
```{r}
factordata %>% finalfit(dependent_selfreferral, explanatory_deprv, metrics = TRUE)
```


Model for self referral vs deprivation + age

```{r}
explanatory_deprv_age <- c ("deprv.factor", "age.factor")
factordata %>% finalfit(dependent_selfreferral, explanatory_deprv_age, metrics = TRUE)
```

Model for self referral vs deprivation + sex

```{r}
explanatory_deprv_sex <- c ("deprv.factor", "sex.factor")
factordata %>% finalfit(dependent_selfreferral, explanatory_deprv_sex, metrics = TRUE)
```

Model for self referral vs deprivation + age + sex

```{r}
explanatory_deprv_sex_age <- c ("deprv.factor", "age.factor", "sex.factor")
factordata %>% finalfit (dependent_selfreferral, explanatory_deprv_sex_age, metrics = TRUE)
```

Model for self referral vs age + sex vs deprv+age+sex
```{r}
explanatory_multi <- c ("age.factor", "sex.factor")
factordata %>% finalfit (dependent_selfreferral, explanatory_deprv_sex_age, explanatory_multi, keep_models = TRUE, metrics = TRUE)
```

Model for self referral vs deprivation +age + sex +age:sex (interaction term)
```{r}
explanatory_allterms <- c ("deprv.factor", "age.factor", "sex.factor", "sex.factor*age.factor")
```
```{r}
factordata %>% finalfit(dependent_selfreferral, explanatory_allterms, metrics = TRUE)
```


Model for self referral vs age + sex + age:sex
```{r}
explanatory_reduced <- c ("age.factor", "sex.factor", "sex.factor*age.factor")
factordata %>% finalfit(dependent_selfreferral, explanatory_reduced, metrics = TRUE)
```
 
 Likelihood ratio test for self referral
```{r}
install.packages("lmtest")
model_fit_self_deprv <- glm(referral_selfreferral ~ deprv.factor, data = factordata, family = binomial)
model_fit_self_allterms <- glm(referral_selfreferral ~ deprv.factor + age.factor + sex.factor + age.factor:sex.factor, data = factordata, family = binomial)
lmtest::lrtest(model_fit_self_deprv, model_fit_self_allterms)
```

Forest plot for self referral
```{r}
tiff("Selfreferral_OR_plot.tiff", width = 8, height = 5, units = 'in', res = 300)
Selfreferral_ORplot <- factordata %>% or_plot(dependent_selfreferral, explanatory_allterms)
dev.off()
```

# Flow Centre Referral

Create summarised column for referral by Flow centre or not
```{r}
factordata <- factordata %>% mutate(referral_flow = fct_collapse(referral.factor,"Not referred by Flow Centre" = c("NHS 24 Referral", "Self-referral", "LUCS Referral", "GP referral", "Police Referral", "Reattend", "999 Emergency", "Misc Referral"))) %>% ff_label ("Flow Centre Referral")
```

Create final fit table for Flow centre referral with p-values displayed. Note: Decile 1 removed as results are being compared relative to it

```{r}
factordata %>% summary_factorlist(dependent   = "referral_flow", explanatory = "deprv.factor", p = TRUE)
```

Create model for flow centre referral vs deprivation + age + sex
```{r}
factordata %>%  summary_factorlist(dependent   = "referral_flow", explanatory = c("deprv.factor", "age.factor", "sex.factor"), p = TRUE)
```


Create model for flow centre referral vs deprivation
```{r}
fit_flow <- glm(referral_flow ~ deprv.factor, data = factordata, family = binomial)
summary(fit_flow)
library(broom)
fit_flow %>% tidy (conf.int = TRUE, exp = TRUE)
```

Note: Desirable model with low AIC and high c-statistic

Model for flow centre referral vs deprivation
```{r}
dependent_flow <- "referral_flow"
explanatory_deprv <- "deprv.factor"
```
```{r}
factordata %>% finalfit(dependent_flow, explanatory_deprv, metrics = TRUE)
```

Model for flow centre referral vs deprivation + age

```{r}
explanatory_deprv_age <- c ("deprv.factor", "age.factor")
factordata %>% finalfit(dependent_flow, explanatory_deprv_age, metrics = TRUE)
```

Model for flow centre referral vs deprivation + sex

```{r}
explanatory_deprv_sex <- c ("deprv.factor", "sex.factor")
factordata %>% finalfit(dependent_flow, explanatory_deprv_sex, metrics = TRUE)
```

Model for flow centre referral vs deprivation + age + sex

```{r}
explanatory_deprv_sex_age <- c ("deprv.factor", "age.factor", "sex.factor")
factordata %>% finalfit (dependent_flow, explanatory_deprv_sex_age, metrics = TRUE)
```

Model for flow centre referral vs age + sex vs deprv+age+sex
```{r}
explanatory_multi <- c ("age.factor", "sex.factor")
factordata %>% finalfit (dependent_flow, explanatory_deprv_sex_age, explanatory_multi, keep_models = TRUE, metrics = TRUE)
```

Model for flow centre referral vs deprivation +age + sex +age:sex (interaction term)
```{r}
explanatory_allterms <- c ("deprv.factor", "age.factor", "sex.factor", "sex.factor*age.factor")
```
```{r}
factordata %>% finalfit(dependent_flow, explanatory_allterms, metrics = TRUE)
```


Model for flow centre referral vs age + sex + age:sex
```{r}
explanatory_reduced <- c ("age.factor", "sex.factor", "sex.factor*age.factor")
factordata %>% finalfit(dependent_flow, explanatory_reduced, metrics = TRUE)
```
 
 Likelihood ratio test for flow centre referral
```{r}
install.packages("lmtest")
model_fit_flow_deprv <- glm(referral_flow ~ deprv.factor, data = factordata, family = binomial)
model_fit_flow_allterms <- glm(referral_flow ~ deprv.factor + age.factor + sex.factor + age.factor:sex.factor, data = factordata, family = binomial)
lmtest::lrtest(model_fit_flow_deprv, model_fit_flow_allterms)
```

Forest plot for flow centre referral
```{r}
tiff("Flowcentre_OR_plot.tiff", width = 8, height = 5, units = 'in', res = 300)
Flowcentre_ORplot <- factordata %>% or_plot(dependent_flow, explanatory_allterms)
dev.off()
```

